\section{Introduction}

\subsection{Background of the Study}
The global financial landscape has experienced unprecedented volatility since 2020, shaped by macroeconomic shocks including the COVID-19 pandemic, shifting monetary policies, and historic inflationary pressures. In this environment, asset classes no longer move in isolation. Investors can no longer rely purely on traditional equity metrics; they must understand how systemic macro indicators such as interest rates, inflation, and market fear—ripple across diverse asset classes simultaneously.

While modern investors increasingly rely on data-driven analysis, financial data remains highly fragmented. Unifying global tech equities, regional banking assets, decentralized cryptocurrencies, and broad market indices into a single analytical framework requires robust data engineering. This study establishes a centralized data pipeline to bridge that gap, allowing for a granular, empirical look at how macroeconomic forces influence a highly diversified digital and traditional asset portfolio.

\subsection{Objectives of the Study}
This study aims to build an integrated financial database by combining daily price data for Apple Inc. (AAPL), Banco de Oro Unibank, Inc. (BDO), Bitcoin (BTC), and the SPDR S\&P 500 ETF Trust (SPY) with selected macroeconomic indicators obtained from the Federal Reserve Economic Data (FRED), namely the 10-Year Treasury Yield, Federal Funds Rate, Consumer Price Index (CPI), and the CBOE Volatility Index (VIX). The datasets cover the period from January 1, 2020 to December 31, 2025 and are integrated using SQL concepts implemented through the \text{dplyr} package in R.

\subsection{Scope of the Analysis}
The integrated database enables the application of data acquisition, cleaning, database integration, descriptive statistical analysis, and visualization techniques to examine the relationships among financial assets and macroeconomic variables. SQL-like operations in R are utilized to organize, merge, and analyze the datasets efficiently, providing meaningful insights into asset performance under different market conditions.

\subsection{Significance of the Study}
The findings of this study are intended to support evidence-based investment decision-making by comparing the performance and risk characteristics of the selected assets. Based on the statistical analyses, SQL investigations, and visualizations, the study concludes with an investment recommendation for allocating a hypothetical portfolio of USD 1,000,000 among the selected assets and cash.

