\section{Data Sources}

Eight datasets were selected for this study, comprising four asset price series and four macroeconomic indicators. All data were obtained from publicly available sources. The sample period from January 1, 2020 to December 31, 2025 was chosen because it encompasses three distinct market regimes: the COVID-19 crash and recovery, the 2021 to 2023 inflation surge, and the subsequent interest rate tightening cycle. This window allows the analysis to capture asset behaviour across crisis, expansion, and transition phases.

The US equity asset is Apple Incorporated (AAPL), listed on the NASDAQ exchange under ticker AAPL. As of 2025, Apple held a market capitalisation of approximately \$4.25 trillion, with a trailing price-to-earnings ratio of 35.08 and an estimated enterprise value of \$4.19 trillion \parencite{stockanalysis2025}. Apple was selected over alternative US technology stocks, particularly NVIDIA, because its valuation and business model present a fundamentally different risk profile that is instructive for a diversified portfolio analysis.

NVIDIA, which briefly overtook Apple as the world's most valuable company in late 2025 and early 2026, trades at a trailing price-to-earnings ratio above 50 and derives over 80\% of its revenue from data centre AI accelerator chips. The concentration of NVIDIA's revenue in a single product category exposed to cyclical capital expenditure cycles and the intense debate among analysts about whether AI infrastructure spending represents a structural transformation or a speculative bubble made it a less suitable choice for a study examining cross-asset behaviour under varying macroeconomic regimes. A stock whose valuation is predominantly determined by a single narrative, however transformative that narrative may be, introduces idiosyncratic risk that complicates the interpretation of portfolio-level results.

Apple presented the opposite profile. Throughout the 2020 to 2025 sample period, Apple was notably deliberate in its approach to artificial intelligence, and in some quarters was characterised as slow to adopt AI relative to its big technology peers. The company did not make early or dramatic public commitments to generative AI, nor did it release large language model products in the manner of its competitors. This restraint was a strategic choice rather than a technological limitation. Apple's approach to AI has consistently prioritised on-device processing and user privacy over cloud-based deployment, a philosophy that the company artefacts under the branding of Apple Intelligence. The June 2026 release of the iOS 19 beta at the Worldwide Developers Conference marked a significant escalation in this strategy \parencite{apple_newsroom2026}, introducing a substantially redesigned Siri powered by Apple Foundation Models that operate through a hybrid on-device and private cloud architecture. This approach, which Apple markets as uniquely privacy-preserving, stands in contrast to the cloud-dependent AI strategies of its competitors and reflects a deliberate product philosophy in which AI is a feature integrated into existing user experiences rather than a standalone product category.

The selection of AAPL over NVIDIA therefore captures a technology sector exposure that is diversified across hardware, services, and increasingly AI-enabled applications, without the single-product concentration risk that characterises the pure-play AI semiconductor firms. This makes AAPL a more representative US equity exposure for a multi-asset portfolio analysis. The Philippine equity asset is BDO Unibank Incorporated (BDO), the largest bank in the Philippines by total assets. As of December 2025, BDO reported total assets of PHP 5.27 trillion, maintaining its ranking as the top universal and commercial bank in the Philippines by asset size \parencite{bsp2025}. BDO was selected to provide exposure to the Philippine financial sector and to emerging market dynamics more broadly, which operate under different monetary policy cycles and economic fundamentals compared to developed markets.

The cryptocurrency asset is Bitcoin (BTC), the largest digital asset by market capitalisation. Bitcoin trades continuously on decentralised cryptocurrency exchanges and was included to evaluate the extent to which its return behaviour correlates with or diverges from traditional asset classes. A secondary objective was to test the frequently asserted claim that Bitcoin functions as an inflation hedge, a proposition examined directly in Q10. The benchmark US equity index is represented by the SPDR S\&P 500 ETF Trust (SPY), which tracks the S\&P 500 index and is one of the most liquid exchange-traded funds globally by assets under management and average daily volume.

Four macroeconomic series were obtained from the Federal Reserve Economic Data (FRED) database maintained by the Federal Reserve Bank of St. Louis. The Consumer Price Index for All Urban Consumers (FRED series ID: CPIAUCSL) measures headline consumer price inflation at monthly frequency and is the most widely referenced US inflation gauge. The 10-Year Treasury Constant Maturity Rate (FRED series ID: DGS10) represents the yield on US government debt at the longest standard tenor and serves as a benchmark for risk-free rates and a barometer of interest rate expectations. The Federal Funds Effective Rate (FRED series ID: FEDFUNDS) captures the overnight rate at which depository institutions lend reserve balances to each other and reflects the stance of US monetary policy. The CBOE Volatility Index (FRED series ID: VIXCLS) tracks the implied volatility of S\&P 500 index options over a 30-day forward period and is commonly interpreted as a gauge of market fear and uncertainty.
