\section{Investment Recommendation}

The results of this research indicate that historical returns are not the only criterion to use when making investment decisions; other factors such as risk, diversification and changing macroeconomic conditions should also be taken into account. Bitcoin had the largest maximum return, the greatest volatility and the greatest downside risk over the sample period of 2020 to 2025 among the selected assets. BDO, on the other hand, had the highest risk-adjusted return and the lowest correlations with all of the other assets, making its greatest contribution to portfolio efficiency rather than maximum absolute returns. Apple has provided significant long-term capital gains with lower volatility compared to Bitcoin, while SPY offered wide exposure to the U.S. equity market and served as the benchmark for the entire analysis. The exploratory analyses and SQL investigations together show that each asset performs differently across different market conditions, underscoring the value of diversification away from investing exclusively in a single high-performing asset.

Based on these findings, this study recommends constructing a diversified portfolio with a good mix of growth and effective risk management. The proposed allocation draws on the analyses of total returns, risk-adjusted performance, cross-asset correlations, tail risk, and macroeconomic sensitivity rather than historical returns alone.

\begin{table}[H]
\centering
\begin{tabular}{lcc}
\toprule
\textbf{Asset} & \textbf{Proposed Allocation} & \textbf{Investment Amount} \\
\midrule
Apple (AAPL) & 35\% & USD 350,000 \\
BDO Unibank & 30\% & USD 300,000 \\
SPDR S\&P 500 ETF Trust (SPY) & 20\% & USD 200,000 \\
Bitcoin (BTC) & 15\% & USD 150,000 \\
\bottomrule
\end{tabular}
\caption{Proposed portfolio allocation.}
\label{tab:allocation}
\end{table}

The proposed allocation is not the result of a quantitative portfolio optimization model, but is meant as a practical allocation suggestion. The allocation is highest for Apple, due to its strong long-term growth and lower risk than Bitcoin. The high risk-adjusted return of BDO and its generally low correlations with the other assets justify a large allocation. SPY offers diversified exposure to the U.S. equity market while helping to minimize the concentration risk of investing in individual securities. Lastly, Bitcoin receives the smallest allocation to account for its outstandingly high returns while limiting the portfolio's exposure to its volatility and tail risk. In summary, the recommended allocation is consistent with the study's key finding: a strategy for balancing return, risk, and diversification is a more sustainable investment approach than the maximization of historical returns.
