---
title: "Appendix: Complete R Code for FINSTAD Case 1"
author: 
  - "CRUZ, Ricardo Miguel Iñigo"
  - "GALEDO, Enrique Lorenzo Hermoso"
  - "SEBALLOS, Josiah Dweyn Panganiban"
  - "SEECHUNG, Camille Castro"
  - "SISON, Aaron Joshua Estacio"
format:
  pdf:
    toc: true
    documentclass: article
    classoption: 12pt
    geometry:
      - margin=1in
    fontsize: 11pt
    code-block-font-size: small
    highlight-style: arrow
    keep-tex: true
include-in-header:
  - text: |
      \usepackage{fvextra}
      \fvset{breaklines=true,breakanywhere=true}
      \RecustomVerbatimEnvironment{Highlighting}{Verbatim}{breaklines=true,breakanywhere=true,commandchars=\\\{\}}
      \let\verbatim\Verbatim
      \let\endverbatim\endVerbatim
      \pagenumbering{arabic}
execute:
  warning: false
  message: false
---

\thispagestyle{empty}
\pagestyle{empty}
# Appendix A1 - Setup & Library Import
Loads the required R libraries (dplyr, tidyr, lubridate, ggplot2, etc.) and defines global helper functions used throughout the analysis.
The working directory is resolved automatically to ensure data file paths work regardless of whether the script runs from the project root or the scripts/ folder.

```{r setup}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(lubridate)
library(DBI)
library(RSQLite)
library(ggplot2)
library(scales)
library(kableExtra)

# Safe table formatter for PDF to prevent overlapping columns and overflowing margins
safe_kable <- function(data, caption = NULL, digits = 4) {
  knitr::kable(data, caption = caption, digits = digits, format = "latex", booktabs = TRUE) %>%
    kable_styling(font_size = 8, latex_options = c("scale_down", "HOLD_position")) %>%
    column_spec(1:ncol(data), width_min = "1.5cm")
}

# Set working directory to project root so data paths resolve consistently
# Ensure data paths resolve correctly regardless of working directory
if (dir.exists("data")) {
  data_dir <- "data"
} else if (dir.exists("../data")) {
  data_dir <- "../data"
} else {
  stop("Cannot find data/ directory. Run from project root or scripts/.")
}

# Global options
options(stringsAsFactors = FALSE, scipen = 999, width = 72)

# Helper to parse BDO volume strings (e.g. "2.50M" to 2500000)
parse_bdo_volume <- function(x) {
  x <- gsub(",", "", x)
  mult <- ifelse(grepl("M", x, ignore.case = TRUE), 1e6,
                 ifelse(grepl("K", x, ignore.case = TRUE), 1e3, 1))
  x_num <- as.numeric(gsub("[A-Za-z%]", "", x))
  x_num * mult
}
```

# Appendix A2 - Ticker Definitions
Defines the asset ticker symbols used throughout the case: AAPL (US equity), BDO.PS (Philippine bank stock), and BTC-USD (cryptocurrency).

```{r data-sources}
# Define tickers
us_asset <- "AAPL"
ph_asset <- "BDO.PS"
crypto <- "BTC-USD"
```

# Appendix A3 - Data Loading
*Author: GALEDO, Enrique Lorenzo Hermoso*
Loads all 8 CSV datasets from the data/ directory into R data frames. Records the original observation count for each dataset before any cleaning operations. Datasets include 4 asset price series (AAPL, BDO, BTC, SPY) and 4 FRED macroeconomic series (CPI, DGS10, FEDFUNDS, VIX). The raw CSVs contain observations well beyond the 2020-2025 analysis window; filtering to the sample period is applied in A4 across all series simultaneously.

```{r data-loading}
# Data Loading -- GALEDO, Enrique Lorenzo Hermoso (Sec 3: Data Acquisition)

# Load all 8 CSVs from the data directory
aapl <- read.csv(file.path(data_dir, "AAPL.csv"), stringsAsFactors = FALSE)
bdo  <- read.csv(file.path(data_dir, "BDO.csv"), stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
btc  <- read.csv(file.path(data_dir, "BTC.csv"), stringsAsFactors = FALSE)
spy  <- read.csv(file.path(data_dir, "SPY.csv"), stringsAsFactors = FALSE)
cpi     <- read.csv(file.path(data_dir, "CPI.csv"), stringsAsFactors = FALSE)
dgs10   <- read.csv(file.path(data_dir, "DGS10.csv"), stringsAsFactors = FALSE)
fedfunds <- read.csv(file.path(data_dir, "FEDFUNDS.csv"), stringsAsFactors = FALSE)
vix     <- read.csv(file.path(data_dir, "VIX.csv"), stringsAsFactors = FALSE)

# Record original observations before any processing
obs_original <- data.frame(
  Dataset = c("AAPL", "BDO", "BTC", "SPY", "CPI", "DGS10", "FEDFUNDS", "VIX"),
  Original_Obs = c(nrow(aapl), nrow(bdo), nrow(btc), nrow(spy),
                   nrow(cpi), nrow(dgs10), nrow(fedfunds), nrow(vix))
)

safe_kable(obs_original, caption="Original Observations per Dataset")
```

# Appendix A4 - Data Cleaning
*Author: SISON, Aaron Joshua Estacio*
Standardises date formats across all 8 datasets, extracts December 2019 seed values for monthly macro series, filters to the 2020-2025 sample period, generates a canonical Mon-Fri business calendar, left-joins each asset onto the calendar, recalculates BTC returns after weekend removal, forward-fills Close prices and zero-fills Daily_Return on non-trading days, seeds CPI and FEDFUNDS with December 2019 values to avoid cold-start nulls, and forward-fills all four macroeconomic variables via LOCF.

```{r data-cleaning}
# Sec 4: Data Cleaning -- SISON, Aaron Joshua Estacio

# Standardise date formats
# AAPL, BTC, SPY: POSIXct-style strings like "2016-06-30 00:00:00-04:00"
aapl <- aapl |> mutate(Date = as.Date(ymd_hms(Date)))
btc  <- btc  |> mutate(Date = as.Date(ymd_hms(Date)))
spy  <- spy  |> mutate(Date = as.Date(ymd_hms(Date)))

# BDO: MM/DD/YYYY
bdo <- bdo |> mutate(Date = mdy(Date))

# CPI, DGS10, FEDFUNDS, VIX: YYYY-MM-DD
cpi     <- cpi     |> mutate(Date = ymd(Date))
dgs10   <- dgs10   |> mutate(Date = ymd(Date))
fedfunds <- fedfunds |> mutate(Date = ymd(Date))
vix     <- vix     |> mutate(Date = ymd(Date))

# Extract December 2019 seed values for CPI and FEDFUNDS
# These provide the initial carry-forward for the first business days
# of January 2020 before the first 2020 monthly release
cpi_dec2019_val <- cpi |> filter(Date == as.Date("2019-12-01")) |> pull(Value)
fedfunds_dec2019_val <- fedfunds |> filter(Date == as.Date("2019-12-01")) |> pull(Value)

# Save pre-filter macro data for CPI_YoY computation (needs 12-month lag)
cpi_pre_filter <- cpi
dgs10_pre_filter <- dgs10
fedfunds_pre_filter <- fedfunds
vix_pre_filter <- vix

# Filter ALL datasets to 2020-01-01 to 2025-12-31
start_date <- as.Date("2020-01-01")
end_date   <- as.Date("2025-12-31")

aapl <- aapl |> filter(Date >= start_date & Date <= end_date)
bdo  <- bdo  |> filter(Date >= start_date & Date <= end_date)
btc  <- btc  |> filter(Date >= start_date & Date <= end_date)
spy  <- spy  |> filter(Date >= start_date & Date <= end_date)
cpi     <- cpi     |> filter(Date >= start_date & Date <= end_date)
dgs10   <- dgs10   |> filter(Date >= start_date & Date <= end_date)
fedfunds <- fedfunds |> filter(Date >= start_date & Date <= end_date)
vix     <- vix     |> filter(Date >= start_date & Date <= end_date)

# Observations after date filtering (before calendar join)
obs_date_filtered <- data.frame(
  Dataset = c("AAPL", "BDO", "BTC", "SPY", "CPI", "DGS10", "FEDFUNDS", "VIX"),
  After_Date_Filter = c(nrow(aapl), nrow(bdo), nrow(btc), nrow(spy),
                        nrow(cpi), nrow(dgs10), nrow(fedfunds), nrow(vix))
)

# Generate canonical Mon-Fri business calendar
# No exchange-specific holiday filtering; holidays are bridged via forward-fill
canonical_calendar <- data.frame(
  Date = seq.Date(from = start_date, to = end_date, by = "day")
) |>
  filter(!weekdays(Date) %in% c("Saturday", "Sunday"))

# Left-join each asset onto the calendar
# Assets with no observation on a given calendar date get NA Close/Daily_Return
# BTC loses approximately 626 weekend rows (2,192 to 1,566)
aapl <- canonical_calendar |> left_join(aapl, by = "Date") |> arrange(Date)
bdo  <- canonical_calendar |> left_join(bdo, by = "Date")  |> arrange(Date)
btc  <- canonical_calendar |> left_join(btc, by = "Date")  |> arrange(Date)
spy  <- canonical_calendar |> left_join(spy, by = "Date")  |> arrange(Date)

# Recalculate daily returns on the filtered business-day grid
# After weekend removal, Monday's return uses Friday's close, not Sunday's
# This must happen AFTER the left-join but BEFORE forward-fill
aapl <- aapl |> arrange(Date) |> mutate(Daily_Return = (Close / lag(Close)) - 1)
bdo  <- bdo  |> arrange(Date) |> mutate(Daily_Return = (Price / lag(Price)) - 1)
btc <- btc |> arrange(Date) |> mutate(Daily_Return = (Close / lag(Close)) - 1)
spy  <- spy  |> arrange(Date) |> mutate(Daily_Return = (Close / lag(Close)) - 1)

# Parse BDO volume strings before column rename
if ("Vol." %in% colnames(bdo)) {
  bdo <- bdo |> mutate(`Vol.` = parse_bdo_volume(`Vol.`))
}

# Rename BDO columns early for forward-fill consistency
bdo <- bdo |> rename(Close = Price, Volume = `Vol.`)

# Seed first-row NAs on 2020-01-01 (New Year's Day holiday, no trading)
aapl_first_close <- aapl |> filter(!is.na(Close)) |> pull(Close) |> head(1)
bdo_first_close <- bdo |> filter(!is.na(Close)) |> pull(Close) |> head(1)
spy_first_close <- spy |> filter(!is.na(Close)) |> pull(Close) |> head(1)
if (is.na(aapl$Close[1])) { aapl$Close[1] <- aapl_first_close }
if (is.na(bdo$Close[1])) { bdo$Close[1] <- bdo_first_close }
if (is.na(spy$Close[1])) { spy$Close[1] <- spy_first_close }

# Forward-fill asset Close prices and zero-fill Daily_Return
# Close: last executable price carried forward (no trade, no price change)
# Daily_Return: zero on non-trading days (no trade, no return)
aapl <- aapl |> tidyr::fill(Close, .direction = "down") |>
  mutate(Daily_Return = if_else(is.na(Daily_Return), 0, Daily_Return))
bdo  <- bdo  |> tidyr::fill(Close, .direction = "down") |>
  mutate(Daily_Return = if_else(is.na(Daily_Return), 0, Daily_Return))
spy  <- spy  |> tidyr::fill(Close, .direction = "down") |>
  mutate(Daily_Return = if_else(is.na(Daily_Return), 0, Daily_Return))
btc  <- btc  |> mutate(Daily_Return = if_else(is.na(Daily_Return), 0, Daily_Return))

# Rename macro variable Value columns to meaningful names
cpi_clean     <- cpi     |> rename(CPI = Value)
dgs10_clean   <- dgs10   |> rename(DGS10 = Value)
fedfunds_clean <- fedfunds |> rename(FEDFUNDS = Value)
vix_clean     <- vix     |> rename(VIX = Value)

# Left-join each macro variable onto the calendar and forward-fill
# CPI: monthly series; seed first row with December 2019 value
cpi <- canonical_calendar |>
  left_join(cpi_clean, by = "Date") |>
  arrange(Date)
cpi_nulls_before <- sum(is.na(cpi$CPI))
if (is.na(cpi$CPI[1])) { cpi$CPI[1] <- cpi_dec2019_val }
cpi <- cpi |> tidyr::fill(CPI, .direction = "down")

# DGS10: business-daily series
# After left-join, gaps exist where FRED did not report
# on days corresponding to non-trading dates in the calendar
dgs10 <- canonical_calendar |>
  left_join(dgs10_clean, by = "Date") |>
  arrange(Date)
dgs10_nulls_before <- sum(is.na(dgs10$DGS10))
dgs10_first <- dgs10 |> filter(!is.na(DGS10)) |> pull(DGS10) |> head(1)
if (is.na(dgs10$DGS10[1])) { dgs10$DGS10[1] <- dgs10_first }
dgs10 <- dgs10 |> tidyr::fill(DGS10, .direction = "down")

# FEDFUNDS: meeting-based series; seed first row with December 2019 value
fedfunds <- canonical_calendar |>
  left_join(fedfunds_clean, by = "Date") |>
  arrange(Date)
fedfunds_nulls_before <- sum(is.na(fedfunds$FEDFUNDS))
if (is.na(fedfunds$FEDFUNDS[1])) { fedfunds$FEDFUNDS[1] <- fedfunds_dec2019_val }
fedfunds <- fedfunds |> tidyr::fill(FEDFUNDS, .direction = "down")

# VIX: business-daily series
vix <- canonical_calendar |>
  left_join(vix_clean, by = "Date") |>
  arrange(Date)
vix_nulls_before <- sum(is.na(vix$VIX))
vix_first <- vix |> filter(!is.na(VIX)) |> pull(VIX) |> head(1)
if (is.na(vix$VIX[1])) { vix$VIX[1] <- vix_first }
vix <- vix |> tidyr::fill(VIX, .direction = "down")

# Observations after calendar join and forward-fill
obs_after_ffill <- data.frame(
  Dataset = c("AAPL", "BDO", "BTC", "SPY", "CPI", "DGS10", "FEDFUNDS", "VIX"),
  After_Calendar = c(nrow(aapl), nrow(bdo), nrow(btc), nrow(spy),
                     nrow(cpi), nrow(dgs10), nrow(fedfunds), nrow(vix)),
  Nulls_Remaining = c(sum(is.na(aapl)), sum(is.na(bdo)), sum(is.na(btc)),
                      sum(is.na(spy)), sum(is.na(cpi)), sum(is.na(dgs10)),
                      sum(is.na(fedfunds)), sum(is.na(vix)))
)

# Forward-fill summary: null counts per macro column before and after
ffill_macro <- data.frame(
  Variable = c("CPI", "DGS10", "FEDFUNDS", "VIX"),
  Nulls_Before_Fill = c(cpi_nulls_before, dgs10_nulls_before,
                        fedfunds_nulls_before, vix_nulls_before),
  Nulls_After_Fill = c(0, 0, 0, 0)
)

# Display tables
safe_kable(obs_date_filtered, caption="Observations After Date Filtering (Before Calendar Join)")
safe_kable(obs_after_ffill, caption="Observations After Calendar Join and Forward-Fill")
safe_kable(ffill_macro, caption="Forward-Fill Summary: Null Counts Before and After")
```

# Appendix A5 - Database Integration
*Author: SISON, Aaron Joshua Estacio*
Demonstrates all four SQL join types (left, inner, full, anti) and justifies the chosen join strategy. Assembles the final analytical dataset by combining asset data with macroeconomic series. Exports the observation reduction table for inclusion in the main report. All macroeconomic variables are forward-filled in A4, so the join does not introduce new null values.

```{r database-integration}
# Sec 5: Database Integration -- SISON, Aaron Joshua Estacio

# Prepare asset data for joining
# BDO columns already renamed in A4 (Price -> Close, Vol. -> Volume)

# Add Asset identifier to each OHLC dataset and select common columns
aapl_clean <- aapl |> mutate(Asset = "AAPL") |>
  select(Date, Asset, Open, High, Low, Close, Volume, Daily_Return)
bdo_clean  <- bdo |> mutate(Asset = "BDO") |>
  select(Date, Asset, Open, High, Low, Close, Volume, Daily_Return)
btc_clean  <- btc |> mutate(Asset = "BTC") |>
  select(Date, Asset, Open, High, Low, Close, Volume, Daily_Return)
spy_clean  <- spy |> mutate(Asset = "SPY") |>
  select(Date, Asset, Open, High, Low, Close, Volume, Daily_Return)

# Combine all assets into one long-format table
assets_all <- bind_rows(aapl_clean, bdo_clean, btc_clean, spy_clean)

# Pivot to wide format: one row per date, columns per asset
assets_wide <- assets_all |>
  select(Date, Asset, Close, Daily_Return) |>
  pivot_wider(names_from = Asset,
              values_from = c(Close, Daily_Return),
              names_sep = "_")

# Macro dataframes from A4 already have renamed columns (CPI, DGS10, FEDFUNDS, VIX)
# and are forward-filled with zero nulls on the canonical calendar

# Demonstration of all four join types

# 1) left_join: keep all asset dates, attach macro data where available
# After forward-fill, left_join produces zero NAs
left_example <- assets_wide |>
  left_join(cpi, by = "Date") |>
  left_join(dgs10, by = "Date") |>
  left_join(fedfunds, by = "Date") |>
  left_join(vix, by = "Date")

# 2) inner_join: only dates present in ALL datasets
# After forward-fill, all series cover all calendar dates
inner_example <- assets_wide |>
  inner_join(cpi, by = "Date") |>
  inner_join(dgs10, by = "Date") |>
  inner_join(fedfunds, by = "Date") |>
  inner_join(vix, by = "Date")

# 3) full_join: all dates from any dataset
full_example <- assets_wide |>
  full_join(cpi, by = "Date") |>
  full_join(dgs10, by = "Date") |>
  full_join(fedfunds, by = "Date") |>
  full_join(vix, by = "Date")

# 4) anti_join: asset dates with NO macro data
# After forward-fill, every date has macro data, so anti_join returns 0 rows
anti_example <- assets_wide |>
  anti_join(cpi, by = "Date")

# Build the final analytical dataset
final_dataset <- left_example

# Observation reduction table
# Original_Obs: row count in each raw CSV (before any processing)
# After_Calendar_Filter: rows remaining after date parsing and 2020-2025 filter
# After_ForwardFill: rows on the canonical calendar after LOCF

obs_original_vec <- obs_original$Original_Obs
names(obs_original_vec) <- obs_original$Dataset

reduction_table <- data.frame(
  Dataset = c("AAPL", "BDO", "BTC", "SPY", "CPI", "DGS10", "FEDFUNDS", "VIX",
              "Assets (wide)", "Final Dataset"),
  Original_Obs = c(obs_original_vec["AAPL"], obs_original_vec["BDO"],
                   obs_original_vec["BTC"], obs_original_vec["SPY"],
                   obs_original_vec["CPI"], obs_original_vec["DGS10"],
                   obs_original_vec["FEDFUNDS"], obs_original_vec["VIX"],
                   NA, NA),
  After_Calendar_Filter = c(nrow(aapl), nrow(bdo), nrow(btc), nrow(spy),
                            nrow(cpi), nrow(dgs10), nrow(fedfunds), nrow(vix),
                            nrow(assets_wide), NA),
  After_ForwardFill = c(NA, NA, NA, NA, nrow(cpi), nrow(dgs10),
                        nrow(fedfunds), nrow(vix), NA, nrow(final_dataset))
)

knitr::kable(reduction_table, caption="Observation Counts Across Processing Stages")
library(knitr)
cat(kable(reduction_table, format = "latex", booktabs = TRUE), file = "../reports/sections/data_cleaning_table.tex")
```

# Appendix A6 - Master Dataset Creation & SQL Database Setup
*Author: SISON, Aaron Joshua Estacio*
Builds the master analytical dataset by restructuring into long format (one row per asset-date combination) and saves it as a persistent SQLite database file with human-readable ISO dates. Pre-calculates CPI year-on-year change for inflation analysis. Exports a CSV copy for compatibility with spreadsheet software. All subsequent SQL questions (A7-A16) query this persisted database.

```{r sql-setup}
# SQLite database setup — creates persisted master dataset

db_path <- "../data/master_dataset.sqlite"
unlink(db_path)
con <- dbConnect(RSQLite::SQLite(), db_path)

# Build long-format analytical dataset for SQL queries
# Each row is one asset-date combination with macro variables attached
sql_long <- bind_rows(
  final_dataset |>
    select(Date, CPI, DGS10, FEDFUNDS, VIX,
           Close = Close_AAPL, Daily_Return = Daily_Return_AAPL) |>
    mutate(Asset = "AAPL"),
  final_dataset |>
    select(Date, CPI, DGS10, FEDFUNDS, VIX,
           Close = Close_BDO, Daily_Return = Daily_Return_BDO) |>
    mutate(Asset = "BDO"),
  final_dataset |>
    select(Date, CPI, DGS10, FEDFUNDS, VIX,
           Close = Close_BTC, Daily_Return = Daily_Return_BTC) |>
    mutate(Asset = "BTC"),
  final_dataset |>
    select(Date, CPI, DGS10, FEDFUNDS, VIX,
           Close = Close_SPY, Daily_Return = Daily_Return_SPY) |>
    mutate(Asset = "SPY")
)

# Pre-calculate CPI YoY for Q10 (inflation analysis)
# CPI is monthly; YoY change = (current / 12-month-ago) - 1
# CPI_YoY is computed from raw monthly CPI values to preserve genuine
# month-over-12-months change, using pre-filter data so the 12-month
# lag resolves correctly for January 2020
cpi_yoy <- cpi_pre_filter |>
  rename(CPI = Value) |>
  arrange(Date) |>
  mutate(CPI_YoY = (CPI / lag(CPI, 12)) - 1) |>
  filter(Date >= start_date & Date <= end_date) |>
  select(Date, CPI_YoY)

# Join CPI_YoY into the long dataset and forward-fill to cover all trading days
sql_long <- sql_long |>
  left_join(cpi_yoy, by = "Date") |>
  arrange(Date) |>
  tidyr::fill(CPI_YoY, .direction = "down")

# Write to SQLite database
dbWriteTable(con, "integrated_data", sql_long, overwrite = TRUE)
fin_data <- tbl(con, "integrated_data")

# Export CSV copy for spreadsheet compatibility (e.g., Microsoft Excel)
write.csv(sql_long, file.path(data_dir, "master_dataset.csv"),
          row.names = FALSE, na = "")

# Show first 5 rows of the master dataset as a preview
nrow_total <- collect(fin_data) |> nrow()
safe_kable(head(sql_long, 5), digits=4, caption=paste0("Master Dataset Preview (First 5 of ", nrow_total, " Rows)"))
```

# Appendix A7 - SQL Question 1: Risk-Adjusted Daily Return
*Author: SISON, Aaron Joshua Estacio*
Computes the ratio of mean daily return to standard deviation for each asset, producing a Sharpe-like measure of return efficiency. The corresponding bar chart is displayed in the Data Visualization section.

```{r q1}
# Q1: Risk-Adjusted Daily Return by Asset -- SISON, Aaron Joshua Estacio

q1 <- fin_data |>
  group_by(Asset) |>
  summarise(
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    SD_Daily_Return = sd(Daily_Return, na.rm = TRUE)
  ) |>
  mutate(Risk_Adjusted_Return = Mean_Daily_Return / SD_Daily_Return) |>
  arrange(desc(Risk_Adjusted_Return)) |>
  collect()

spy_ratio <- q1 |> filter(Asset == "SPY") |> pull(Risk_Adjusted_Return)

safe_kable(q1, digits=4, caption="Q1 Data Output")
```

# Appendix A8 - SQL Question 2: COVID-19 Crash Cumulative Returns
*Author: SISON, Aaron Joshua Estacio*
Measures crisis resilience during the February to March 2020 COVID-19 market crash by computing cumulative returns for each asset.

```{r q2}
# Q2: COVID-19 Crash Cumulative Returns -- SISON, Aaron Joshua Estacio

q2_raw <- fin_data |>
  select(Date, Asset, Daily_Return) |>
  collect() |>
  mutate(Date = as.Date(Date))

q2 <- q2_raw |>
  filter(Date >= as.Date("2020-02-01") & Date <= as.Date("2020-03-31")) |>
  filter(!is.na(Daily_Return)) |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(Cumulative_Return = cumprod(1 + Daily_Return) - 1)

q2_summary <- q2 |>
  summarise(
    Start_Date = min(Date),
    End_Date = max(Date),
    Total_Cumulative_Return = last(Cumulative_Return),
    Observations = n()
  )

safe_kable(q2_summary, digits=4, caption="Q2 Summary Data Output")
```

# Appendix A9 - SQL Question 3: Pairwise Correlation of Daily Returns
*Author: SISON, Aaron Joshua Estacio*
Calculates the Pearson correlation coefficients for daily returns across all asset pairs and displays the correlation matrix.

```{r q3}
# Q3: Pairwise Correlation of Daily Returns -- SISON, Aaron Joshua Estacio

returns_wide <- fin_data |>
  select(Date, Asset, Daily_Return) |>
  collect() |>
  pivot_wider(names_from = Asset, values_from = Daily_Return)

cor_matrix <- returns_wide |>
  select(AAPL, BDO, BTC, SPY) |>
  cor(use = "pairwise.complete.obs")

cor_long <- as.data.frame(cor_matrix) |>
  tibble::rownames_to_column("Asset1") |>
  tidyr::pivot_longer(-Asset1, names_to = "Asset2", values_to = "Correlation") |>
  mutate(Asset1 = factor(Asset1, levels = rev(c("AAPL", "BDO", "BTC", "SPY"))),
         Asset2 = factor(Asset2, levels = c("AAPL", "BDO", "BTC", "SPY")))

safe_kable(round(cor_matrix, 4), caption="Q3 Correlation Matrix Output")
```

# Appendix A10 - SQL Question 4: Asset Ranking by Total Return
*Author: CRUZ, Ricardo Miguel Iñigo*
Ranks assets by total price return and annualized return over the full 2020-2025 sample period.

```{r q4}
# Q4: Asset Ranking by Total Return -- CRUZ, Ricardo Miguel Inigo

q4_raw <- fin_data |>
  select(Date, Asset, Close, Daily_Return) |>
  collect() |>
  mutate(Date = as.Date(Date))

q4 <- q4_raw |>
  filter(!is.na(Close), !is.na(Daily_Return)) |>
  group_by(Asset) |>
  arrange(Date) |>
  summarise(
    Start_Date = min(Date),
    End_Date = max(Date),
    Start_Close = first(Close),
    End_Close = last(Close),
    Total_Return = (End_Close / Start_Close) - 1,
    Annualized_Return = (1 + Total_Return)^(365.25 / as.numeric(End_Date - Start_Date)) - 1,
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    SD_Daily_Return = sd(Daily_Return, na.rm = TRUE),
    Observations = n(),
    .groups = "drop"
  ) |>
  arrange(desc(Total_Return))

q4_plot <- q4 |>
  mutate(Asset = factor(Asset, levels = q4$Asset[order(q4$Total_Return)]))

safe_kable(q4, digits=4, caption="Q4 Data Output")
```

# Appendix A11 - SQL Question 5: Best and Worst Trading Days
*Author: CRUZ, Ricardo Miguel Iñigo*
Identifies extreme single-day gains and losses for each asset to assess tail risk.

```{r q5}
# Q5: Best and Worst Trading Days -- CRUZ, Ricardo Miguel Inigo

q5_raw <- fin_data |>
  select(Date, Asset, Daily_Return) |>
  collect() |>
  mutate(Date = as.Date(Date))

q5 <- q5_raw |>
  filter(!is.na(Daily_Return)) |>
  group_by(Asset) |>
  summarise(
    Best_Date = Date[which.max(Daily_Return)],
    Best_Daily_Return = max(Daily_Return, na.rm = TRUE),
    Worst_Date = Date[which.min(Daily_Return)],
    Worst_Daily_Return = min(Daily_Return, na.rm = TRUE),
    Return_Range = Best_Daily_Return - Worst_Daily_Return,
    Observations = n(),
    .groups = "drop"
  ) |>
  arrange(Worst_Daily_Return)

safe_kable(q5, digits=4, caption="Q5 Data Output")
```

# Appendix A12 - SQL Question 6: Moving Average Crossover Analysis
*Author: CRUZ, Ricardo Miguel Iñigo*
Implements a 20-day and 60-day simple moving average crossover strategy to analyze momentum regime shifts.

```{r q6}
# Q6: Moving Average Crossover Analysis -- CRUZ, Ricardo Miguel Inigo

q6_raw <- fin_data |>
  select(Date, Asset, Close, Daily_Return) |>
  collect() |>
  mutate(Date = as.Date(Date))

q6_ma <- q6_raw |>
  filter(!is.na(Close)) |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(
    MA_20 = as.numeric(stats::filter(Close, rep(1 / 20, 20), sides = 1)),
    MA_60 = as.numeric(stats::filter(Close, rep(1 / 60, 60), sides = 1)),
    Signal = case_when(
      MA_20 > MA_60 ~ "Bullish",
      MA_20 < MA_60 ~ "Bearish",
      TRUE ~ "Neutral"
    )
  ) |>
  ungroup()

q6 <- q6_ma |>
  filter(!is.na(MA_20), !is.na(MA_60), !is.na(Daily_Return)) |>
  group_by(Asset, Signal) |>
  summarise(
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    SD_Daily_Return = sd(Daily_Return, na.rm = TRUE),
    Observations = n(),
    .groups = "drop"
  ) |>
  arrange(Asset, desc(Mean_Daily_Return))

q6_latest_signal <- q6_ma |>
  filter(!is.na(MA_20), !is.na(MA_60)) |>
  group_by(Asset) |>
  slice_max(Date, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(Asset, Date, Close, MA_20, MA_60, Signal) |>
  arrange(Asset)

safe_kable(q6, digits=4, caption="Q6 Data Output")
```

# Appendix A13 - SQL Question 7: Volume and Volatility Relationship
*Author: CRUZ, Ricardo Miguel Iñigo*
Tests whether elevated trading volume systematically coincides with higher volatility by partitioning days into high-volume and normal-volume regimes.

```{r q7}
# Q7: Volume and Volatility Relationship -- CRUZ, Ricardo Miguel Inigo

volume_lookup <- assets_all |>
  select(Date, Asset, Volume) |>
  mutate(
    Date = as.Date(Date),
    Volume = as.numeric(Volume)
  )

q7_raw <- fin_data |>
  select(Date, Asset, Daily_Return) |>
  collect() |>
  mutate(
    Date = as.Date(Date),
    Daily_Return = as.numeric(Daily_Return)
  ) |>
  left_join(volume_lookup, by = c("Date", "Asset")) |>
  filter(
    !is.na(Daily_Return),
    !is.na(Volume),
    Volume > 0
  )

q7 <- q7_raw |>
  group_by(Asset) |>
  mutate(
    Volume_Threshold = quantile(Volume, 0.75, na.rm = TRUE),
    Volume_Regime = if_else(
      Volume >= Volume_Threshold,
      "High Volume Days",
      "Normal Volume Days"
    ),
    Absolute_Return = abs(Daily_Return)
  ) |>
  ungroup() |>
  group_by(Asset, Volume_Regime) |>
  summarise(
    Observations = n(),
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    Mean_Absolute_Return = mean(Absolute_Return, na.rm = TRUE),
    Volatility = sd(Daily_Return, na.rm = TRUE),
    Average_Volume = mean(Volume, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(Asset, Volume_Regime)

safe_kable(q7, digits=4, caption="Q7 Data Output")
```

# Appendix A14 - SQL Question 8: VIX Regime Analysis
*Author: SISON, Aaron Joshua Estacio*
Assesses asset performance during periods of high versus low market fear by segregating trading days based on VIX levels.

```{r q8}
# Q8: VIX Regime Analysis -- SISON, Aaron Joshua Estacio

q8_raw <- fin_data |>
  filter(!is.na(VIX)) |>
  select(Asset, Date, Daily_Return, VIX) |>
  collect() |>
  mutate(Date = as.Date(Date))

q8 <- q8_raw |>
  mutate(
    VIX_Regime = case_when(
      VIX > 30 ~ "High Volatility (VIX > 30)",
      VIX < 20 ~ "Low Volatility (VIX < 20)",
      TRUE ~ "Moderate (20 <= VIX <= 30)"
    )
  ) |>
  group_by(Asset, VIX_Regime) |>
  summarise(
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    SD_Daily_Return = sd(Daily_Return, na.rm = TRUE),
    Observations = n(),
    .groups = "drop"
  ) |>
  arrange(Asset, VIX_Regime)

q8_plot <- q8 |>
  filter(VIX_Regime %in% c("High Volatility (VIX > 30)", "Low Volatility (VIX < 20)")) |>
  mutate(VIX_Regime = recode(VIX_Regime,
    "High Volatility (VIX > 30)" = "VIX > 30 (High Fear)",
    "Low Volatility (VIX < 20)" = "VIX < 20 (Low Fear)"))

safe_kable(q8, digits=4, caption="Q8 Data Output")
```

# Appendix A15 - SQL Question 9: Interest Rate Sensitivity
*Author: SISON, Aaron Joshua Estacio*
Measures asset sensitivity to changes in the 10-year Treasury yield by comparing mean returns on days when yields rise versus fall.

```{r q9}
# Q9: Interest Rate Sensitivity -- SISON, Aaron Joshua Estacio

q9_raw <- fin_data |>
  filter(!is.na(DGS10)) |>
  select(Asset, Date, Daily_Return, DGS10) |>
  collect() |>
  mutate(Date = as.Date(Date))

q9 <- q9_raw |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(DGS10_Change = DGS10 - lag(DGS10)) |>
  ungroup() |>
  filter(!is.na(DGS10_Change)) |>
  mutate(
    Yield_Direction = case_when(
      DGS10_Change > 0 ~ "Yields Rising",
      DGS10_Change < 0 ~ "Yields Falling",
      TRUE ~ "No Change"
    )
  ) |>
  group_by(Asset, Yield_Direction) |>
  summarise(
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    SD_Daily_Return = sd(Daily_Return, na.rm = TRUE),
    Observations = n(),
    .groups = "drop"
  ) |>
  arrange(Asset, Yield_Direction)

q9_plot <- q9 |>
  filter(Yield_Direction %in% c("Yields Rising", "Yields Falling"))

safe_kable(q9, digits=4, caption="Q9 Data Output")
```

# Appendix A16 - SQL Question 10: Inflation Hedge Performance
*Author: SISON, Aaron Joshua Estacio*
Evaluates which assets served as effective inflation hedges during the June 2021 to February 2023 high-inflation period (CPI YoY above 5%).

```{r q10}
# Q10: Inflation Hedge Performance -- SISON, Aaron Joshua Estacio

q10_raw <- fin_data |>
  select(Date, Asset, Daily_Return, CPI_YoY) |>
  collect() |>
  mutate(Date = as.Date(Date))

q10 <- q10_raw |>
  filter(!is.na(CPI_YoY) & CPI_YoY > 0.05) |>
  filter(!is.na(Daily_Return)) |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(Cumulative_Return = cumprod(1 + Daily_Return) - 1) |>
  summarise(
    Start_Date = min(Date),
    End_Date = max(Date),
    Start_CPI_YoY = first(CPI_YoY),
    End_CPI_YoY = last(CPI_YoY),
    Total_Cumulative_Return = last(Cumulative_Return),
    Mean_Daily_Return = mean(Daily_Return, na.rm = TRUE),
    Observations = n()
  ) |>
  arrange(desc(Total_Cumulative_Return))

q10_plot <- q10 |>
  mutate(Asset = factor(Asset, levels = q10$Asset[order(q10$Total_Cumulative_Return)]))

safe_kable(q10, digits=4, caption="Q10 Data Output")
```

# Appendix A17 - Data Visualization Charts
*Author: SEBALLOS, Josiah Dweyn Panganiban*
Contains the ggplot code for all 10 figures corresponding to SQL Questions Q1 through Q10.
Each figure is saved as a PDF to the reports/figures/ directory for inclusion in the main report's Data Visualization section.

```{r dir-setup}
# Create figures directory if it does not exist
dir.create("../reports/figures", recursive = TRUE, showWarnings = FALSE)
```

### Figure 1: Risk-Adjusted Daily Return by Asset
Bar chart showing the ratio of mean daily return to standard deviation for each asset. The dashed reference line marks SPY's ratio.

```{r viz-q1}
ggplot(q1, aes(x = Risk_Adjusted_Return, y = reorder(Asset, Risk_Adjusted_Return),
                fill = Asset)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_vline(xintercept = spy_ratio, linetype = "dashed",
             color = "#B3B3B3", linewidth = 0.4) +
  geom_text(aes(x = Risk_Adjusted_Return / 2,
                label = sprintf("%.4f", Risk_Adjusted_Return)),
            hjust = 0.5, size = 3.2, color = "white", fontface = "bold") +
  annotate("text", x = spy_ratio, y = 4.5, label = "SPY benchmark",
           size = 2.8, color = "#595959", hjust = -0.1) +
  scale_fill_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                BTC = "#F7931A", SPY = "#00733E")) +
  labs(title = "Risk-Adjusted Daily Return by Asset",
       subtitle = "Mean return divided by standard deviation, 2020 to 2025",
       x = "Risk-Adjusted Return", y = NULL) +
  xlim(0, 0.07) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_q1.pdf", width = 7, height = 3.5, device = "pdf")
```

### Figure 2: Cumulative Returns During the COVID-19 Crash
Line chart tracking each asset's cumulative return from February to March 2020, indexed to 100% at period start.

```{r viz-q2}
ggplot(q2, aes(x = Date, y = Cumulative_Return, color = Asset)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#B3B3B3", linewidth = 0.3) +
  scale_color_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                 BTC = "#F7931A", SPY = "#00733E")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Cumulative Returns During the COVID-19 Crash",
       subtitle = "February to March 2020, indexed to 100% at period start",
       x = "Date", y = "Cumulative Return", color = "Asset") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "bottom")
ggsave("../reports/figures/fig_q2.pdf", width = 7, height = 4, device = "pdf")
```

### Figure 3: Pairwise Correlation Heatmap
Heatmap matrix of Pearson correlation coefficients for daily returns across the four assets.

```{r viz-q3}
ggplot(cor_long, aes(x = Asset2, y = Asset1, fill = Correlation)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = sprintf("%.3f", Correlation)), size = 3.5) +
  scale_fill_gradient2(low = "#1DC9A4", mid = "#F2F2F2", high = "#2E45B8",
                       midpoint = 0.4, limits = c(0, 1)) +
  labs(title = "Pairwise Correlation of Daily Returns",
       subtitle = "Pearson correlation coefficients, 2020 to 2025",
       x = NULL, y = NULL, fill = "Correlation") +
  theme_minimal(base_size = 11) +
  theme(panel.grid = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "right")
ggsave("../reports/figures/fig_q3.pdf", width = 6, height = 4.5, device = "pdf")
```

### Figure 4: Asset Ranking by Total Return
Horizontal bar chart ranking assets by total percentage price return from 2020 to 2025.

```{r viz-q4}
ggplot(q4_plot, aes(x = Total_Return, y = Asset, fill = Asset)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_vline(xintercept = 0, linewidth = 0.3, color = "#595959") +
  geom_text(aes(x = Total_Return / 2,
                label = sprintf("%+.1f%%", Total_Return * 100)),
            hjust = 0.5, size = 3.2, color = "white", fontface = "bold") +
  scale_fill_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                BTC = "#F7931A", SPY = "#00733E")) +
  scale_x_continuous(labels = scales::percent_format(),
                     expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = "Asset Ranking by Total Return",
       subtitle = "Total price return from 2020 to 2025",
       x = "Total Return", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_q4.pdf", width = 7, height = 3.5, device = "pdf")
```

### Figure 5: Best and Worst Single-Day Returns
Grouped bar chart comparing the largest single-day gain and loss for each asset.

```{r viz-q5}
q5_plot <- q5 |>
  select(Asset, Best_Daily_Return, Worst_Daily_Return) |>
  pivot_longer(cols = c(Best_Daily_Return, Worst_Daily_Return),
               names_to = "Trading_Day_Type",
               values_to = "Daily_Return") |>
  mutate(
    Trading_Day_Type = recode(Trading_Day_Type,
                              Best_Daily_Return = "Best Trading Day",
                              Worst_Daily_Return = "Worst Trading Day")
  )

ggplot(q5_plot, aes(x = Asset, y = Daily_Return, fill = Trading_Day_Type)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = 0, linewidth = 0.3, color = "#595959") +
  scale_fill_manual(values = c("Best Trading Day" = "#2E45B8",
                                "Worst Trading Day" = "#C91D42")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Best and Worst Single-Day Returns",
       subtitle = "Largest one-day gains and losses by asset, 2020 to 2025",
       x = NULL, y = "Daily Return", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "bottom")
ggsave("../reports/figures/fig_q5.pdf", width = 7, height = 4, device = "pdf")
```

### Figure 6: 20-Day and 60-Day Moving Average Crossover
Multi-panel faceted line chart showing closing prices with 20-day and 60-day moving averages from 2024 to 2025.

```{r viz-q6}
q6_plot <- q6_ma |>
  filter(Date >= as.Date("2024-01-01"))

ggplot(q6_plot, aes(x = Date)) +
  geom_line(aes(y = Close), linewidth = 0.4, color = "#595959") +
  geom_line(aes(y = MA_20), linewidth = 0.5, color = "#2E45B8") +
  geom_line(aes(y = MA_60), linewidth = 0.5, color = "#C91D42") +
  facet_wrap(~ Asset, scales = "free_y") +
  labs(title = "20-Day and 60-Day Moving Average Crossover",
       subtitle = "Bullish signal when 20-day MA is above 60-day MA, 2024 to 2025",
       x = NULL, y = "Closing Price") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_q6.pdf", width = 7, height = 4.5, device = "pdf")
```

### Figure 7: Volume and Volatility Relationship
Grouped bar chart comparing mean absolute daily return during high-volume versus normal-volume days.

```{r viz-q7}
ggplot(q7, aes(x = Asset, y = Mean_Absolute_Return, fill = Volume_Regime)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Volume and Volatility Relationship",
    subtitle = "Mean absolute daily return during high-volume versus normal-volume days, 2020 to 2025",
    x = NULL,
    y = "Mean Absolute Daily Return",
    fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(size = 9, color = "#595959"),
    legend.position = "bottom"
  )
ggsave("../reports/figures/fig_q7.pdf", width = 7, height = 4, device = "pdf")
```

### Figure 8: Mean Daily Return by VIX Regime
Grouped bar chart comparing asset returns during high-fear (VIX > 30) versus low-fear (VIX < 20) regimes.

```{r viz-q8}
ggplot(q8_plot, aes(x = Asset, y = Mean_Daily_Return, fill = VIX_Regime)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = 0, linewidth = 0.3, color = "#595959") +
  scale_fill_manual(values = c("VIX > 30 (High Fear)" = "#C91D42",
                                "VIX < 20 (Low Fear)" = "#2E45B8")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Mean Daily Return by VIX Regime",
       subtitle = "Comparing asset performance under high versus low market fear, 2020 to 2025",
       x = NULL, y = "Mean Daily Return", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "bottom")
ggsave("../reports/figures/fig_q8.pdf", width = 7, height = 4, device = "pdf")
```

### Figure 9: Mean Daily Return by 10-Year Treasury Yield Direction
Grouped bar chart showing asset returns conditional on whether daily yields are rising or falling.

```{r viz-q9}
ggplot(q9_plot, aes(x = Asset, y = Mean_Daily_Return, fill = Yield_Direction)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = 0, linewidth = 0.3, color = "#595959") +
  scale_fill_manual(values = c("Yields Rising" = "#F97A1F",
                                "Yields Falling" = "#2E45B8")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Mean Daily Return by 10-Year Treasury Yield Direction",
       subtitle = "Asset returns conditional on daily yield changes, 2020 to 2025",
       x = NULL, y = "Mean Daily Return", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "bottom")
ggsave("../reports/figures/fig_q9.pdf", width = 7, height = 4, device = "pdf")
```

### Figure 10: Cumulative Return During High Inflation Period
Horizontal bar chart showing total cumulative returns during the period when CPI year-on-year exceeded 5%.

```{r viz-q10}
ggplot(q10_plot, aes(x = Total_Cumulative_Return, y = Asset, fill = Asset)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(x = Total_Cumulative_Return / 2,
                label = sprintf("%+.1f%%", Total_Cumulative_Return * 100)),
            hjust = 0.5, size = 3.2, color = "white", fontface = "bold") +
  scale_fill_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                BTC = "#F7931A", SPY = "#00733E")) +
  scale_x_continuous(labels = scales::percent_format(),
                     expand = expansion(mult = c(0.15, 0.15))) +
  labs(title = "Cumulative Return During High Inflation Period",
       subtitle = "June 2021 to February 2023, CPI year-on-year above 5%",
       x = "Cumulative Return", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_q10.pdf", width = 7, height = 3.5, device = "pdf")
```

# Appendix A18 - Exploratory Data Visualization
*Author: SEBALLOS, Josiah Dweyn Panganiban*
Contains 8 exploratory figures covering cumulative returns, price trends, return distributions, rolling volatility, drawdown analysis, risk-return profiles, and portfolio risk contribution. These complement the SQL-driven charts in A17.

### Viz Figure 1: Cumulative Returns by Asset
Tracks the growth of USD 1 invested in each asset from 2020 to 2025, indexed to 100% at the start.

```{r viz-j1}
fig1_data <- sql_long |>
  filter(!is.na(Daily_Return)) |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(Cumulative_Return = cumprod(1 + Daily_Return) - 1) |>
  ungroup()

ggplot(fig1_data, aes(x = Date, y = Cumulative_Return, color = Asset)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#B3B3B3", linewidth = 0.3) +
  scale_color_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                 BTC = "#F7931A", SPY = "#00733E")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Cumulative Returns by Asset",
       subtitle = "Growth of $1 invested, 2020 to 2025",
       x = "Date", y = "Cumulative Return", color = "Asset") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "bottom")
ggsave("../reports/figures/fig_viz1.pdf", width = 7, height = 4, device = "pdf")
```


### Viz Figure 2: Closing Price Trends by Asset
Plots nominal closing prices on individual scales to compare price-level dynamics across assets of different magnitudes.

```{r viz-j2}
ggplot(sql_long, aes(x = Date, y = Close, color = Asset)) +
  geom_line(linewidth = 0.7, show.legend = FALSE, na.rm = TRUE) +
  facet_wrap(~ Asset, scales = "free_y", ncol = 2) +
  scale_color_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                BTC = "#F7931A", SPY = "#00733E")) +
  labs(title = "Closing Price Trends by Asset",
       subtitle = "2020 to 2025",
       x = "Date", y = "Closing Price") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_viz2.pdf", width = 7, height = 5, device = "pdf")
```


### Viz Figure 3: Return Distribution with Skewness and Kurtosis
Overlays each asset's daily return histogram against a fitted normal curve to assess distribution shape and tail risk.

```{r viz-j3}
fig3_data <- sql_long |> filter(!is.na(Daily_Return))

calc_skew <- function(x) {
  n <- length(x); m <- mean(x)
  (sum((x - m)^3) / n) / (sd(x)^3)
}
calc_kurt <- function(x) {
  n <- length(x); m <- mean(x)
  (sum((x - m)^4) / n) / (sd(x)^2)^2 - 3
}

fig3_stats <- fig3_data |>
  group_by(Asset) |>
  summarise(
    Mean = mean(Daily_Return), SD = sd(Daily_Return),
    Skewness = calc_skew(Daily_Return), Kurtosis = calc_kurt(Daily_Return),
    .groups = "drop"
  ) |>
  mutate(label = sprintf("Skew: %.2f | Kurt: %.2f", Skewness, Kurtosis))

normal_curves <- fig3_stats |>
  rowwise() |>
  reframe(
    Asset = Asset,
    x = seq(Mean - 4 * SD, Mean + 4 * SD, length.out = 200),
    y = dnorm(x, mean = Mean, sd = SD)
  )

ggplot(fig3_data, aes(x = Daily_Return)) +
  geom_histogram(aes(y = after_stat(density), fill = Asset), bins = 60,
                  alpha = 0.75, show.legend = FALSE) +
  geom_line(data = normal_curves, aes(x = x, y = y), color = "black", linewidth = 0.5) +
  geom_text(data = fig3_stats, aes(x = -Inf, y = Inf, label = label),
            hjust = -0.05, vjust = 1.5, size = 2.8) +
  facet_wrap(~ Asset, scales = "free", ncol = 2) +
  scale_fill_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                BTC = "#F7931A", SPY = "#00733E")) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Return Distribution with Skewness & Kurtosis",
       subtitle = "Daily returns vs. fitted normal distribution (black line), 2020 to 2025",
       x = "Daily Return", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_viz3.pdf", width = 7, height = 5, device = "pdf")
```


### Viz Figure 4: Return Range and Spread by Asset
Boxplot showing median, interquartile range, and outlier days for each asset.

```{r viz-j4}
ggplot(fig3_data, aes(x = Asset, y = Daily_Return, fill = Asset)) +
  geom_boxplot(show.legend = FALSE, outlier.size = 0.8, outlier.alpha = 0.4) +
  scale_fill_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                BTC = "#F7931A", SPY = "#00733E")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Return Range & Spread by Asset",
       subtitle = "Median, interquartile range, and outlier days, 2020 to 2025",
       x = NULL, y = "Daily Return") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_viz4.pdf", width = 7, height = 4, device = "pdf")
```


### Viz Figure 5: 30-Day Rolling Volatility by Asset
Tracks rolling 30-day standard deviation of daily returns to visualize volatility evolution over time.

```{r viz-j5}
rolling_sd <- function(x, n = 30) {
  sapply(seq_along(x), function(i) {
    if (i < n) return(NA)
    sd(x[(i - n + 1):i], na.rm = TRUE)
  })
}

fig5_data <- sql_long |>
  filter(!is.na(Daily_Return)) |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(Rolling_Vol = rolling_sd(Daily_Return, 30)) |>
  ungroup() |>
  filter(!is.na(Rolling_Vol))

ggplot(fig5_data, aes(x = Date, y = Rolling_Vol, color = Asset)) +
  geom_line(linewidth = 0.7) +
  scale_color_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                 BTC = "#F7931A", SPY = "#00733E")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "30-Day Rolling Volatility by Asset",
       subtitle = "Rolling standard deviation of daily returns, 2020 to 2025",
       x = "Date", y = "30-Day Rolling Volatility", color = "Asset") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"),
        legend.position = "bottom")
ggsave("../reports/figures/fig_viz5.pdf", width = 7, height = 4, device = "pdf")
```


### Viz Figure 6: Drawdown from Peak by Asset
Computes percentage decline from each asset's running historical high to visualize recovery patterns.

```{r viz-j6}
fig6_data <- fig1_data |>
  group_by(Asset) |>
  arrange(Date) |>
  mutate(Running_Max = cummax(1 + Cumulative_Return),
         Drawdown = (1 + Cumulative_Return) / Running_Max - 1) |>
  ungroup()

ggplot(fig6_data, aes(x = Date, y = Drawdown, color = Asset)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ Asset, ncol = 2) +
  scale_color_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                 BTC = "#F7931A", SPY = "#00733E")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Drawdown from Peak by Asset",
       subtitle = "Percentage decline from running maximum value, 2020 to 2025",
       x = "Date", y = "Drawdown") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_viz6.pdf", width = 7, height = 5, device = "pdf")
```


### Viz Figure 7: Risk vs. Return by Asset
Scatter plot comparing mean daily return against standard deviation across all four assets.

```{r viz-j7}
fig7_data <- sql_long |>
  filter(!is.na(Daily_Return)) |>
  group_by(Asset) |>
  summarise(Mean_Return = mean(Daily_Return), SD_Return = sd(Daily_Return), .groups = "drop")

ggplot(fig7_data, aes(x = SD_Return, y = Mean_Return, color = Asset)) +
  geom_point(size = 4) +
  geom_text(aes(label = Asset), vjust = -1.2, size = 3.5, show.legend = FALSE) +
  scale_color_manual(values = c(AAPL = "#555555", BDO = "#003D6B",
                                 BTC = "#F7931A", SPY = "#00733E")) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Risk vs. Return by Asset",
       subtitle = "Mean daily return vs. standard deviation, 2020 to 2025",
       x = "Risk (Std. Dev. of Daily Return)", y = "Mean Daily Return") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_viz7.pdf", width = 7, height = 5, device = "pdf")
```


### Viz Figure 8: Equal-Weighted Portfolio Risk Contribution
Decomposes total portfolio variance into each asset's contribution to show that equal dollar weights do not produce equal risk.

```{r viz-j8}
returns_matrix <- sql_long |>
  filter(Asset %in% c("AAPL", "BDO", "BTC"), !is.na(Daily_Return)) |>
  select(Date, Asset, Daily_Return) |>
  pivot_wider(names_from = Asset, values_from = Daily_Return) |>
  select(AAPL, BDO, BTC) |>
  na.omit()

cov_matrix <- cov(returns_matrix)
weights <- c(AAPL = 1/3, BDO = 1/3, BTC = 1/3)
port_var <- as.numeric(t(weights) %*% cov_matrix %*% weights)
marginal_contrib <- cov_matrix %*% weights
risk_contrib <- as.numeric(weights * marginal_contrib / port_var)

fig8_data <- data.frame(Asset = names(weights), Risk_Contribution = risk_contrib)

ggplot(fig8_data, aes(x = Asset, y = Risk_Contribution, fill = Asset)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(Risk_Contribution, accuracy = 0.1)),
            vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c(AAPL = "#555555", BDO = "#003D6B", BTC = "#F7931A")) +
  scale_y_continuous(labels = scales::percent_format(), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Equal-Weighted Portfolio Risk Contribution",
       subtitle = "Share of total portfolio variance contributed by each asset",
       x = NULL, y = "Share of Portfolio Risk") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, color = "#595959"))
ggsave("../reports/figures/fig_viz8.pdf", width = 7, height = 4, device = "pdf")
```


# Appendix A19 - Descriptive Statistics
*Author: GALEDO, Enrique Lorenzo Hermoso*
*Note: Corresponds to Section 6 (Descriptive and Visualizations). Code is being developed in a separate PR and will be merged here.*

```{r descriptive-stats}
# To be merged from GALEDO, Enrique Lorenzo Hermoso
```

# Appendix A20 - Investment Recommendation
*Author: SEECHUNG, Camille Castro*
*Note: Corresponds to Section 7 (Investment Recommendation). Allocate USD 1,000,000.*

```{r investment-rec}
# Sec 7: Investment Recommendation -- USD 1,000,000 allocation among AAPL, BDO, BTC, cash -- SEECHUNG, Camille Castro
```

# Appendix A21 - Executive Dashboard
*Author: CRUZ, Ricardo Miguel Iñigo*
*Note: Corresponds to Section 0 (Executive Summary).*

```{r dashboard-prep}
# Sec 0: Executive Dashboard -- key stats, charts, SQL findings, recommendation summary -- CRUZ, Ricardo Miguel Iñigo
```
