---
title: "FINSTAD Case 1: Building an Integrated Financial Database"
author: 
  - "CRUZ, Ricardo Miguel Iñigo"
  - "GALEDO, Enrique Lorenzo Hermoso"
  - "SEBALLOS, Josiah Dweyn Panganiban"
  - "SEECHUNG, Camille Castro"
  - "SISON, Aaron Joshua Estacio"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: cosmo
execute:
  warning: false
  message: false
---

# 1. Introduction

```{r setup}
# Load necessary libraries
library(tidyverse)
library(lubridate)
library(quantmod)
library(DBI)
library(RSQLite)
```

# 2. Data Sources
*Notes on selected U.S. asset, PH asset, crypto, SPY, and FRED variables.*

```{r data-sources}
# Define tickers
us_asset <- ""
ph_asset <- ""
crypto <- ""
```

# 3. Data Acquisition
*Acquire data, identify source, sample period, original observations, missing values, duplicates.*

```{r data-acquisition}
# Fetch data code goes here
```

# 4. Data Cleaning
*Handle missing values, duplicates, return calculations.*

```{r data-cleaning}
# Cleaning code goes here
```

# 5. Database Integration
*Perform joins and justify strategy.*

```{r database-integration}
# Merging code goes here (left_join, inner_join, etc.)
```

# 6. SQL-Based Financial Investigation
*10 meaningful financial questions using select, filter, mutate, arrange, group_by, summarise.*

```{r sql-investigation}
# Set up in-memory SQLite database for dplyr/SQL translation
con <- dbConnect(RSQLite::SQLite(), ":memory:")
# Write integrated dataset to db
# dbWriteTable(con, "financial_data", final_dataset)

# Query 1
# ...
```

# 7. Descriptive Statistical Analysis
*Mean, median, variance, std dev, min, max, range, skewness, kurtosis.*

```{r descriptive-stats}
# Stats code goes here
```

# 8. Data Visualization
*At least 8 professional-quality figures.*

```{r data-viz-1}
# ggplot code goes here
```

# 9. Exploratory Financial Analysis
*Compare returns, volatility, inflationary periods, high-VIX periods, interest-rate environments.*

```{r exploratory-analysis}
# Analysis code goes here
```

# 10. Investment Recommendation
*Allocate USD 1,000,000.*

```{r investment-rec}
# Portfolio allocation calculations
```

# 11. Executive Dashboard (Data Prep)

```{r dashboard-prep}
# Prep data to be exported or charted for the 1-page dashboard
```
